% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}

\documentclass[]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi

% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
  \usepackage{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
  \usepackage[left=2.5in,bottom=1.25in,top=1.25in,right=1in]{geometry}
  \ifxetex
\usepackage[setpagesize=false, % page size defined by xetex
            unicode=false, % unicode breaks when used with xetex
            xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
bookmarks=true,
pdfauthor={},
pdftitle={Expaining Atheism - Pilot 1: CREDs/CRUDs},
colorlinks=true,
citecolor=blue,
urlcolor=blue,
linkcolor=magenta,
pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

% Customization for cos_prereg
\usepackage{longtable,booktabs,threeparttable,tabularx}
\linespread{1.5}
\newcounter{question}
\setcounter{question}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\setlength{\droptitle}{-2em}
\title{Expaining Atheism - Pilot 1: CREDs/CRUDs}
\pretitle{\begin{changemargin}{-8pc}{0pc} \centering\large Preregistration\\ \Huge}
\posttitle{\end{changemargin}}
  \author{
          Connair J. S. Russell\textsuperscript{1},
          Jon Lanman\textsuperscript{1},
          Aiyana Willard\textsuperscript{2},
          Adam Baimel\textsuperscript{3}          \\ \vspace{0.5cm}
              \textsuperscript{1} Queen's University Belfast\\
              \textsuperscript{2} Brunel University London\\
              \textsuperscript{3} Oxford Brookes University      }

  \def\affdep{{"", "", "", ""}}%
  \def\affcity{{"", "", "", ""}}%
  \preauthor{\begin{changemargin}{-8pc}{0pc} \centering\large}
  \postauthor{\end{changemargin}}
\date{INSERT DATE}
\predate{\begin{changemargin}{-8pc}{0pc} \centering\large\emph}
\postdate{\end{changemargin}}
\usepackage{fancyhdr}
\pagestyle{fancy}
\setlength{\headheight}{14.0pt}
\renewcommand{\headrulewidth}{0pt}
\lhead{}
\rhead{\large\textsc{\MakeLowercase{EA Prereg 1}}}



% Title settings
\usepackage{titlesec}
\titleformat{\section}[display]{\bfseries\Large}{\thesection}{}{}[]
\titlespacing{\section}{0pc}{*3}{*1.5}
\titleformat{\subsection}[leftmargin]{\titlerule\bfseries\filleft}{\thesubsection}{.5em}{}
\titlespacing{\subsection}{8pc}{5ex plus .1ex minus .2ex}{1.5pc}


% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi


\begin{document}
\maketitle
\vspace{2pc}


\ifPDFTeX
  \newcommand\Question[2]{%
   \leavevmode\par
   \stepcounter{question}
   \noindent
   \textbf{\thequestion. #1}. #2\par}
\else % if luatex or xetex
  \NewCommandCopy{\oldQuestion}{\Question}
  \renewcommand\Question[2]{%
   \leavevmode\par
   \stepcounter{question}
   \noindent
   \textbf{\thequestion. #1}. #2\par}
\fi



\newcommand\Answer[1]{%
    \noindent
    \textit{Registered response}: #1\par}

\hypertarget{study-information}{%
\section{Study Information}\label{study-information}}

\hypertarget{title}{%
\subsection{Title}\label{title}}

EA Preregistration: Cognitive Style and Non-verbal Reasoning

\hypertarget{description}{%
\subsection{Description}\label{description}}

This preregistration is part of a number of preregistrations for the
Explaining Atheism project. Our other registrations can be found on the
OSF here {[}ADD LINK{]}, with copies on our FigShare, and on my GitHub.

This registration is for the testing of the measurement and effect of
Credibility Enhancing Displays (CREDs), and Credibility Undermining
Displays (CRUDs) on atheism/non-belief. Specifically, this element of
our first pilot project will test the scales used for the measurement of
CREDs and CRUDs, and the relationship between these measures and
measures of non-belief.

Credibility Enhancing Displays are costly displays made by a model,
particularly those that reflect a statement made by that model. Displays
that show a model walking the walk, or practicing what they preach, if
you will. For example if a model says a blue mushroom is toxic and then
avoids eating it even when hungry, or priestly celibacy would be CREDS.
A number of studies have shown historical CRED exposure to predict
religiosity (CITES). The present project looks to build upon this, and
also update the measures for cross-cultural comparison. Conversely CRUDs
are behaviours which undermine credibility. They are behaviours where a
model might engage in behaviours which contradict their stated belief,
such as breaking celibacy, or eating haram foods.

\hypertarget{hypotheses}{%
\subsection{Hypotheses}\label{hypotheses}}

\textbf{CREDs}

Lower historic CRED exposure will predict lower:

\begin{enumerate}
\def\labelenumi{(\roman{enumi})}
\tightlist
\item
  Religious Identification
\item
  Belief in God
\end{enumerate}

and greater:

\begin{enumerate}
\def\labelenumi{(\roman{enumi})}
\tightlist
\item
  Agnosticism
\end{enumerate}

\textbf{CRUDs}

Greater historic CRUD exposure will predict lower:

\begin{enumerate}
\def\labelenumi{(\roman{enumi})}
\tightlist
\item
  Religious Identification
\item
  Belief in God
\end{enumerate}

and greater:

\begin{enumerate}
\def\labelenumi{(\roman{enumi})}
\tightlist
\item
  Atheist identification
\end{enumerate}

\hypertarget{design-plan}{%
\section{Design Plan}\label{design-plan}}

\hypertarget{study-type}{%
\subsection{Study type}\label{study-type}}

\textbf{Observational Study}. Data is collected from study subjects that
are not randomly assigned to a treatment. This includes surveys, natural
experiments, and regression discontinuity designs.

\hypertarget{study-design}{%
\subsection{Study design}\label{study-design}}

This registration is for the analysis of data collected as part of our
first pilot which is comprised of 5 separate surveys, one for each of
the explanatory variable clusters noted in the main pre-registration
(see here). This specific registration concerns our socialisation
survey, and is run in the in the UK, Japan and Brazil.

\hypertarget{randomization}{%
\subsection{Randomization}\label{randomization}}

Enter your response here.

\hypertarget{sampling-plan}{%
\section{Sampling Plan}\label{sampling-plan}}

We will collect representative samples for each nation. This will be
done by an independent survey company, and the samples will be
representative for the national population by Age, Sex and Geographic
Region.

\hypertarget{existing-data}{%
\subsection{Existing data}\label{existing-data}}

\textbf{Registration prior to creation of data}. As of the date of
submission of this research plan for preregistration, the data have not
yet been collected, created, or realized.

\hypertarget{data-collection-procedures}{%
\subsection{Data collection
procedures}\label{data-collection-procedures}}

Participants are recruited by the independent survey company, Savanta
Comres, from their pool of participants. Participants will be paid a
local rate for their participation, and will be selected to meet
representativeness requirements outlined above.

\hypertarget{sample-size}{%
\subsection{Sample size}\label{sample-size}}

The sample size is 500 per nation.

Brazil = 500

UK = 500

Japan = 500

\hypertarget{sample-size-rationale}{%
\subsection{Sample size rationale}\label{sample-size-rationale}}

Samples are the maximally viable sample for the financial means of the
project.

\hypertarget{variables}{%
\section{Variables}\label{variables}}

\hypertarget{measured-variables}{%
\subsection{Measured variables}\label{measured-variables}}

The explanatory variables of primary interest, CRUDs and CREDs are part
of a survey with the following explanatory variables. See attached
variables document for definitions:

\begin{itemize}
\item
  CREDs

  \begin{itemize}
  \item
    CRED Score
  \item
    CRED Score (Parental)
  \item
    CRED Score (Group)
  \item
    CRED Score (Group Leader)
  \item
    CRED Score (Group Members)
  \end{itemize}
\item
  CRUDs

  \begin{itemize}
  \item
    CRUD Score
  \item
    CRUD Score (Parental)
  \item
    CRUD Score (Group)
  \item
    CRUD Score (Group Leader)
  \item
    CRUD Score (Group Members)
  \end{itemize}
\item
  Non-theistic Socialisation
\item
  Normativity of Religion:

  \begin{itemize}
  \item
    Normativity of Religion
  \item
    Normativity (Nation)
  \item
    Normativity (Ethnicity)
  \item
    Normativity (Peers)
  \item
    Normativity (Group)
  \end{itemize}
\item
  Religious Emphasis
\end{itemize}

Similarly the outcome variables of interest are part of a wide range of
response variables that are collected, see the attached codebook for all
other variables. The outcome variables by which we will test our
hypotheses are as follows. Again, see attached variables document for
detailed defintions:

\begin{itemize}
\tightlist
\item
  Belief in God (continuous)
\item
  Belief in God (binary)
\item
  Religious Identification
\item
  Agnosticism
\item
  Atheist Identification
\end{itemize}

Details of what comprises each of these measures and scales can be found
in the attached codebook.

\hypertarget{indices}{%
\subsection{Indices}\label{indices}}

See attached codebook.

\hypertarget{analysis-plan}{%
\section{Analysis Plan}\label{analysis-plan}}

We will run confirmatory analyses for each of our response variables
noted above variables, and all the explanatory variables above. That is,
we will have the following analyses:

\emph{Confirmatory Analyses}

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Analysis & Explanatory Variables & Dependent Variable \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
CRED 1 & CRED Score & Belief in God (Continuous) \\
CRED 2 & CRED Score & Belief in God (Binary) \\
CRED 3 & CRED Score & Religious Identification \\
CRED 5 & CRED Score & Agnosticism \\
CRED 5 & CRED Score & Atheist Identification \\
CRUD 1 & CRUD Score & Belief in God (Continuous) \\
CRUD 2 & CRUD Score & Belief in God (Binary) \\
CRUD 3 & CRUD Score & Religious Identification \\
CRUD 5 & CRUD Score & Agnosticism \\
CRUD 5 & CRUD Score & Atheist Identification \\
\end{longtable}

\hypertarget{statistical-models}{%
\subsection{Statistical models}\label{statistical-models}}

Bayesian Regression Models will be used to make inferences in the first
instance.

\hypertarget{transformations}{%
\subsection{Transformations}\label{transformations}}

Both Atheist Identification, and Belief in God (binary) dummy coded
variables of a different question in the surveys. For details see the
attached variables and codebook files.

\hypertarget{inference-criteria}{%
\subsection{Inference criteria}\label{inference-criteria}}

Given that we will be using Bayesian statistical models, our inference
criteria will be the 89\% credible intervals for parameter values, in
the first instance.

\hypertarget{data-exclusion}{%
\subsection{Data exclusion}\label{data-exclusion}}

An attention check is included in data collection, however participants
who fail this are excluded from the data at data collection phase and
replaced with a new participant, as such will not be in the dataset.

\hypertarget{missing-data}{%
\subsection{Missing data}\label{missing-data}}

The data collection process does not allow missing data.

\hypertarget{exploratory-analyses-optional}{%
\subsection{Exploratory analyses
(optional)}\label{exploratory-analyses-optional}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1090}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.7115}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1795}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Socialisation 1 & CRUD Score, CRUD Score, Normativity of Religion,
Non-theistic Socialisation, Religious Emphasis & Belief in God
(Continuous) \\
Socialisation 2 & CRUD Score, CRUD Score, Normativity of Religion,
Non-theistic Socialisation, Religious Emphasis & Belief in God
(Binary) \\
Socialisation 3 & CRUD Score, CRUD Score, Normativity of Religion,
Non-theistic Socialisation, Religious Emphasis & Religious
Identification \\
Socialisation 5 & CRUD Score, CRUD Score, Normativity of Religion,
Non-theistic Socialisation, Religious Emphasis & Agnosticism \\
Socialisation 5 & CRUD Score, CRUD Score, Normativity of Religion,
Non-theistic Socialisation, Religious Emphasis & Atheist
Identification \\
CRED 1 & CRED Score, CRED Score (Parental), CRED Score (Group), Cred
Score (Group Leaders), CRED Score (Group Members) & Belief in God
(Continuous) \\
CRED 2 & CRED Score, CRED Score (Parental), CRED Score (Group), Cred
Score (Group Leaders), CRED Score (Group Members) & Belief in God
(Binary) \\
CRED 3 & CRED Score, CRED Score (Parental), CRED Score (Group), Cred
Score (Group Leaders), CRED Score (Group Members) & Religious
Identification \\
CRED 5 & CRED Score, CRED Score (Parental), CRED Score (Group), Cred
Score (Group Leaders), CRED Score (Group Members) & Agnosticism \\
CRED 5 & CRED Score, CRED Score (Parental), CRED Score (Group), Cred
Score (Group Leaders), CRED Score (Group Members) & Atheist
Identification \\
\end{longtable}

We will also run analyses of the psychometric properties of the CRED and
CRUD scales, along with measurement invariance analyses.

\hypertarget{other}{%
\section{Other}\label{other}}

\hypertarget{other-optional}{%
\subsection{Other (Optional)}\label{other-optional}}

This work is part of a larger scale project using the same data. See
decription for further information.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{section}{%
\subsection{}\label{section}}

\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}

\noindent

\end{document}
