---
title: "Expaining Atheism Preregistration: Pilot 1, Cognitive Style and Non-verbal Reasoning"
author:
- name: Connair J. S. Russell
  affiliation: 1
- name: Jon Lanman
  affiliation: 1
- name: Aiyana Willard
  affiliation: 2
- name: Adam Baimel
  affiliation: 3
date: "INSERT DATE"
output:
  word_document: default
  pdf_document: prereg::prereg_pdf
  html_document: default
affiliation:
- id: 1
  institution: Queen's University Belfast
- id: 2
  institution: Brunel University London
- id: 3
  institution: Oxford Brookes University
shorttitle: EA Prereg 1
---

# Study Information

## Title

```{=html}
<!-- Provide the working title of your study. It may be the same title that you submit for publication of your final manuscript, but it is not a requirement. The title should be a specific and informative description of a project. Vague titles such as 'Fruit fly preregistration plan' are not appropriate.

Example: Effect of sugar on brownie tastiness. -->
```
EA Preregistration Pilot 1: Cognitive Style & Non-Verbal Reasoning

## Description

```{=html}
<!-- Please give a brief description of your study, including some background, the purpose of the of the study, or broad research questions. The description should be no longer than the length of an abstract. It can give some context for the proposed study, but great detail is not needed here for your preregistration.

Example: Though there is strong evidence to suggest that sugar affects taste preferences, the effect has never been demonstrated in brownies. Therefore, we will measure taste preference for four different levels of sugar concentration in a standard brownie recipe to determine if the effect exists in this pastry. -->
```
This preregistration is part of a number of preregistrations for the Explaining Atheism project. Our other registrations can be found on the OSF here [ADD LINK], with copies on our FigShare, and on my GitHub.

This registration is for the testing of the effects of Thinking/Cognitive Styles and Non-Verbal Reasoning on atheism/non-belief. Both cognitive ability and an analytic cognitive style have been argued to be predictors of non-belief or atheism, however mixed evidence exists for both. We will use a non-verbal reasoning task, often used as a measure of cognitive ability, to conduct a large scale cross-cultural examination of the effect of non-verbal reasoning on levels of belief. Similarly we will use an established thinking styles measures to test its effect on belief.

## Hypotheses

```{=html}
<!-- List specific, concise, and testable hypotheses. Please state if the hypotheses are directional or non-directional. If directional, state the direction. A predicted effect is also appropriate here. If a specific interaction or moderation is important to your research, you can list that as a separate hypothesis.

Example: If taste affects preference, then mean preference indices will be higher with higher concentrations of sugar. -->
```
Greater analytic thinking scores will predict lower levels of belief.

Greater scores on non-verbal reasoning will predict lower levels of belief.

# Design Plan

<!-- In this section, you will be asked to describe the overall design of your study. Remember that this research plan is designed to register a single study, so if you have multiple experimental designs, please complete a separate preregistration. -->

## Study type

**Observational Study**. Data is collected from study subjects that are not randomly assigned to a treatment. This includes surveys, natural experiments, and regression discontinuity designs.

## Study design

```{=html}
<!-- Describe your study design. Examples include two-group, factorial, randomized block, and repeated measures. Is it a between (unpaired), within-subject (paired), or mixed design? Describe any counterbalancing required. Typical study designs for observation studies include cohort, cross sectional, and case-control studies.

This question has a variety of possible answers. The key is for a researcher to be as detailed as is necessary given the specifics of their design. Be careful to determine if every parameter has been specified in the description of the study design. There may be some overlap between this question and the following questions. That is OK, as long as sufficient detail is given in one of the areas to provide all of the requested information. For example, if the study design describes a complete factorial, 2 X 3 design and the treatments and levels are specified previously, you do not have to repeat that information.

Example: We have a between subjects design with 1 factor (sugar by mass) with 4 levels. -->
```
This registration is for the analysis of data collected as part of our first pilot which is comprised of 5 separate surveys, one for each of the explanatory variable clusters noted in the exploratory pre-registration (see here: ). This specific registration concerns our "Thinking styles" explanatory variables, and is run in the in the UK and Brazil..

## Randomization

```{=html}
<!-- If you are doing a randomized study, how will you randomize, and at what level? Typical randomization techniques include: simple, block, stratified, and adaptive covariate randomization. If randomization is required for the study, the method should be specified here, not simply the source of random numbers.

Example: We will use block randomization, where each participant will be randomly assigned to one of the four equally sized, predetermined blocks. The random number list used to create these four blocks will be created using the web applications available at https://random.org. -->
```
The order of the items within both the Non-Verbal Reasoning and the Cognitive Styles questionnaires were randomised. Both the General Belief and Supernatural Belief questionnaires were also randomised within questionnaire.

# Sampling Plan

<!-- In this section we’ll ask you to describe how you plan to collect samples, as well as the number of samples you plan to collect and your rationale for this decision. Please keep in mind that the data described in this section should be the actual data used for analysis, so if you are using a subset of a larger dataset, please describe the subset that will actually be used in your study. -->

We will collect representative samples for each nation. This will be done by an independent survey company, and the samples will be representative for the national population by Age, Sex and Geographic Region.

## Existing data

<!-- Preregistration is designed to make clear the distinction between confirmatory tests, specified prior to seeing the data, and exploratory analyses conducted after observing the data. Therefore, creating a research plan in which existing data will be used presents unique challenges. Please select the description that best describes your situation. Please do not hesitate to contact us if you have questions about how to answer this question (prereg@cos.io). -->

**Registration prior to accessing the data**. As of the date of submission, the data exist, but have not been accessed by you or your collaborators. Commonly, this includes data that has been collected by another researcher or institution.

At the time of registration the data is being collected by the survey company on our teams behalf. No member of our research team has seen any of the data, nor any summary statistics.

## Data collection procedures

```{=html}
<!-- Please describe the process by which you will collect your data. If you are using human subjects, this should include the population from which you obtain subjects, recruitment efforts, payment for participation, how subjects will be selected for eligibility from the initial pool (e.g. inclusion and exclusion rules), and your study timeline. For studies that donÍt include human subjects, include information about how you will collect samples, duration of data gathering efforts, source or location of samples, or batch numbers you will use.

The answer to this question requires a specific set of instructions so that another person could repeat the data collection procedures and recreate the study population. Alternatively, if the study population would be unable to be reproduced because it relies on a specific set of circumstances unlikely to be recreated (e.g., a community of people from a specific time and location), the criteria and methods for creating the group and the rationale for this unique set of subjects should be clear.

Example: Participants will be recruited through advertisements at local pastry shops. Participants will be paid $10 for agreeing to participate (raised to $30 if our sample size is not reached within 15 days of beginning recruitment). Participants must be at least 18 years old and be able to eat the ingredients of the pastries. -->
```
Participants are recruited by the independent survey company, Savanta Comres, from their pool of participants. Participants will be paid a local rate for their participation, and will be selected to meet representativeness requirements outlined above.

## Sample size

```{=html}
<!-- Describe the sample size of your study. How many units will be analyzed in the study? This could be the number of people, birds, classrooms, plots, interactions, or countries included. If the units are not individuals, then describe the size requirements for each unit. If you are using a clustered or multilevel design, how many units are you collecting at each level of the analysis? For some studies, this will simply be the number of samples or the number of clusters. For others, this could be an expected range, minimum, or maximum number.

Example: Our target sample size is 280 participants. We will attempt to recruit up to 320, assuming that not all will complete the total task. -->
```
The sample size is 500 participants per nation.

Brazil = 500

UK = 500

## Sample size rationale

```{=html}
<!-- This could include a power analysis or an arbitrary constraint such as time, money, or personnel. This gives you an opportunity to specifically state how the sample size will be determined. A wide range of possible answers is acceptable; remember that transparency is more important than principled justifications. If you state any reason for a sample size upfront, it is better than stating no reason and leaving the reader to "fill in the blanks." Acceptable rationales include: a power analysis, an arbitrary number of subjects, or a number based on time or monetary constraints.

Example: We used the software program G*Power to conduct a power analysis. Our goal was to obtain .95 power to detect a medium effect size of .25 at the standard .05 alpha error probability. -->
```
Samples are the maximally viable sample for the financial means of the project.

# Variables

<!-- In this section you can describe all variables (both manipulated and measured variables) that will later be used in your confirmatory analysis plan. In your analysis plan, you will have the opportunity to describe how each variable will be used. If you have variables which you are measuring for exploratory analyses, you are not required to list them, though you are permitted to do so. -->

## Measured variables

```{=html}
<!-- Describe each variable that you will measure. This will include outcome measures, as well as any predictors or covariates that you will measure. You do not need to include any variables that you plan on collecting if they are not going to be included in the confirmatory analyses of this study.

Observational studies and meta-analyses will include only measured variables. As with the previous questions, the answers here must be precise. For example, 'intelligence,' 'accuracy,' 'aggression,' and 'color' are too vague. Acceptable alternatives could be 'IQ as measured by Wechsler Adult Intelligence Scale' 'percent correct,' 'number of threat displays,' and 'percent reflectance at 400 nm.'

Example: The single outcome variable will be the perceived tastiness of the single brownie each participant will eat. We will measure this by asking participants ‘How much did you enjoy eating the brownie’ (on a scale of 1-7, 1 being 'not at all', 7 being 'a great deal') and 'How good did the brownie taste' (on a scale of 1-7, 1 being 'very bad', 7 being 'very good'). -->
```
The explanatory variables of primary interest, Non-Verbal Reasoning and Cognitive Style are part of a survey with the following explanatory variables. See attached variables document for definitions:

-   Non-Verbal Reasoning
-   Cognitve Style (correct)

Similarly the outcome variables of interest are part of a wide range of response variables that are collected, see the attached codebook for all other variables. The outcome variables by which we will test our hypotheses are as follows. Again, see attached variables document for detailed defintions:

-   Belief in God (continuous)
-   Belief in God (binary)
-   Religious Identification

Details of what comprises each of these measures and scales can be found in the attached codebook.

## Indices

```{=html}
<!-- If any measurements are  going to be combined into an index (or even a mean), what measures will you use and how will they be combined? Include either a formula or a precise description of your method. If your are using a more complicated statistical method to combine measures (e.g. a factor analysis), you can note that here but describe the exact method in the analysis plan section.

If you are using multiple pieces of data to construct a single variable, how will this occur? Both the data that are included and the formula or weights for each measure must be specified. Standard summary statistics, such as "means" do not require a formula, though more complicated indices require either the exact formula or, if it is an established index in the field, the index must be unambiguously defined. For example, "biodiversity index" is too broad, whereas "Shannon’s biodiversity index" is appropriate.

Example: We will take the mean of the two questions above to create a single measure of 'brownie enjoyment.'  -->
```
See attached codebook.

# Analysis Plan

```{=html}
<!-- You may describe one or more confirmatory analysis in this preregistration. Please remember that all analyses specified below must be reported in the final article, and any additional analyses must be noted as exploratory or hypothesis generating.

A confirmatory analysis plan must state up front which variables are predictors (independent) and which are the outcomes (dependent), otherwise it is an exploratory analysis. You are allowed to describe any exploratory work here, but a clear confirmatory analysis is required. -->
```
We will run confirmatory analyses for each of our response variables noted above, for both of the two explanatory variables. We will also run a number of exploratory analyses outlined in the "Exploratory Analyses" section below. The confirmatory analyses are as follows:

*Confirmatory Analyses*

| Analysis             | Explanatory Variables     | Outcome Variable           |
|----------------------|---------------------------|----------------------------|
| Cognitive Style 1    | Cognitive Style (correct) | Belief in God (Continuous) |
| Cognitive Style 2    | Cognitive Style (correct) | Belief in God (Binary)     |
| Cognitive Style 3    | Cognitive Style (correct) | Religious Identification   |
| Non-Verbal Reasoning | Non-Verbal Reasoning      | Belief in God (Continuous) |
| Non-Verbal Reasoning | Non-Verbal Reasoning      | Belief in God (Binary)     |
| Non-Verbal Reasoning | Non-Verbal Reasoning      | Religious Identification   |

## Statistical models

```{=html}
<!-- What statistical model will you use to test each hypothesis? Please include the type of model (e.g. ANOVA, multiple regression, SEM, etc) and the specification of the model (this includes each variable that will be included as predictors, outcomes, or covariates). Please specify any interactions, subgroup analyses, pairwise or complex contrasts, or follow-up tests from omnibus tests. If you plan on using any positive controls, negative controls, or manipulation checks you may mention that here. Remember that any test not included here must be noted as an exploratory test in your final article.

This is perhaps the most important and most complicated question within the preregistration. As with all of the other questions, the key is to provide a specific recipe for analyzing the collected data. Ask yourself: is enough detail provided to run the same analysis again with the information provided by the user? Be aware for instances where the statistical models appear specific, but actually leave openings for the precise test. See the following examples:

- If someone specifies a 2x3 ANOVA with both factors within subjects, there is still flexibility with the various types of ANOVAs that could be run. Either a repeated measures ANOVA (RMANOVA) or a multivariate ANOVA (MANOVA) could be used for that design, which are two different tests. 
- If you are going to perform a sequential analysis and check after 50, 100, and 150 samples, you must also specify the p-values you’ll test against at those three points.

Example:  We will use a one-way between subjects ANOVA to analyze our results. The manipulated, categorical independent variable is 'sugar' whereas the dependent variable is our taste index. -->
```
Bayesian Regression Models will be used to make inferences in the first instance.

Model specifications can be found in the attached Excel file.

We will treat outcome variables that are single Likert scale measures as ordinal, and as such will use cumulative probit regression following the following structure:

```{r, eval=FALSE}

library("brms")

exampleProbitModel <-
  brm(formula = outcome_variable ~ explanatory_variable_1 +
                                   explanatory_variable_2 +
                                   explanatory_variable_n,
      prior = priors,
      data = data, 
      family = cumulative(link = "probit"),
      chains = 4,
      cores = 4,
      threads = 2,
      iter = 3000,
      warmup = 1000,
      seed = 2023, 
      backend = "cmdstanr")

```

For binary outcomes we will use logistic regression with the following format:

```{r, eval=FALSE}

library("brms")

exampleLogisticModel<- 
  brm(formula = outcome_variable ~ explanatory_variable_1 +
                                   explanatory_variable_2 +
                                   explanatory_variable_n,
      prior = priors,
      data = data, 
      family = bernoulli(link = "logit"),
      chains = 4,
      cores = 4,
      threads = 2,
      iter = 3000,
      warmup = 1000,
      seed = 2023, 
      backend = "cmdstanr")

```

And for continuous outcomes we will use gaussian multiple regression:

```{r, eval=FALSE}

library("brms")

exampleLogisticModel<- 
  brm(formula = outcome_variable ~ explanatory_variable_1 +
                                   explanatory_variable_2 +
                                   explanatory_variable_n,
      prior = priors,
      data = data, 
      family = gaussian(),
      chains = 4,
      cores = 4,
      threads = 2,
      iter = 3000,
      warmup = 1000,
      seed = 2023, 
      backend = "cmdstanr")

```

We will set the following uninformative priors for our outcome variables:

For betas and intercepts we will set a normally distributed prior with a mean of 0 and a standard deviation of 1.

For ordinal probit models we will set a normal distribution for the thresholds, following [Kurz](https://solomonkurz.netlify.app/blog/2021-12-29-notes-on-the-bayesian-cumulative-probit/).

## Transformations

```{=html}
<!-- If you plan on transforming, centering, recoding the data, or will require a coding scheme for categorical variables, please describe that process. If any categorical predictors are included in a regression, indicate how those variables will be coded (e.g. dummy coding, summation coding, etc.) and what the reference category will be.

Example: The "Effect of sugar on brownie tastiness" does not require any additional transformations. However, if it were using a regression analysis and each level of sweet had been categorically described (e.g. not sweet, somewhat sweet, sweet, and very sweet), 'sweet' could be dummy coded with 'not sweet' as the reference category. -->
```
In some instances response variables are dummy coded binary variables from a categorical response option - see attached codebook for all variable definitions.

As a default we will center likert response scores at the center of the scale, unless the scale has a meaningful interval (e.g. age, n years).

## Inference criteria

```{=html}
<!-- What criteria will you use to make inferences? Please describe the information youÍll use (e.g. p-values, bayes factors, specific model fit indices), as well as cut-off criterion, where appropriate. Will you be using one or two tailed tests for each of your analyses? If you are comparing multiple conditions or testing multiple hypotheses, will you account for this?
 
p-values, confidence intervals, and effect sizes are standard means for making an inference, and any level is acceptable, though some criteria must be specified in this or previous fields. Bayesian analyses should specify a Bayes factor or a credible interval. If you are selecting models, then how will you determine the relative quality of each? In regards to multiple comparisons, this is a question with few "wrong" answers. In other words, transparency is more important than any specific method of controlling the false discovery rate or false error rate. One may state an intention to report all tests conducted or one may conduct a specific correction procedure; either strategy is acceptable.

Example: We will use the standard p<.05 criteria for determining if the ANOVA and the post hoc test suggest that the results are significantly different from those expected if the null hypothesis were correct. The post-hoc Tukey-Kramer test adjusts for multiple comparisons. -->
```
In the first instance, inferences about whether a variable has or does not have a theoretically relevant effect will be done based on examination of the posterior distributions of parameter estimates.

More specifically we will examine the extent to which the credible intervals of these posterior distributions overlap with a Region of Practical Equivalence (ROPE), that is, a specified region we consider to be theoretically equivalent to zero. This will be done using the {bayestestR} package (<https://easystats.github.io/bayestestR/articles/region_of_practical_equivalence.html>) in R. For our Gaussian and Cumulative Probit models we will use the default range -0.1 to 0.1 for standardized effect sizes, which equates to a "negligible effect size", and -0.18 to 0.18 for logistic models.

## Data exclusion

```{=html}
<!-- How will you determine what data or samples, if any, to exclude from your analyses? How will outliers be handled? Will you use any awareness check? Any rule for excluding a particular set of data is acceptable. One may describe rules for excluding a participant or for identifying outlier data.

Example: No checks will be performed to determine eligibility for inclusion besides verification that each subject answered each of the three tastiness indices. Outliers will be included in the analysis. -->
```
Participants who fail data quality checks are removed and replaced by the survey provider prior to our analysis. In the first instance outliers will not be removed, however sensitivity analysis will be run on any published final models.

## Missing data

```{=html}
<!-- How will you deal with incomplete or missing data? Any relevant explanation is acceptable. As a final reminder, remember that the final analysis must follow the specified plan, and deviations must be either strongly justified or included as a separate, exploratory analysis.

Example: If a subject does not complete any of the three indices of tastiness, that subject will not be included in the analysis. -->
```
The data collection process does not allow missing data.

## Exploratory analyses (optional)

```{=html}
<!-- If you plan to explore your data set to look for unexpected differences or relationships, you may describe those tests here. An exploratory test is any test where a prediction is not made up front, or there are multiple possible tests that you are going to use. A statistically significant finding in an exploratory test is a great way to form a new confirmatory hypothesis, which could be registered at a later time.

Example: We expect that certain demographic traits may be related to taste preferences. Therefore, we will look for relationships between demographic variables (age, gender, income, and marital status) and the primary outcome measures of taste preferences. -->
```
We will run measurement invariance analyses to see to what extent the measures used vary across cultures.

# Other

## Other (Optional)

<!-- If there is any additional information that you feel needs to be included in your preregistration, please enter it here. Literature cited, disclosures of any related work such as replications or work that uses the same data, or other context that will be helpful for future readers would be appropriate here. -->

This work is part of a larger scale project using the same data. See decription for further information.

# References

## 

```{=tex}
\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}
```
\noindent
